\begin{romanpages}      % roman-numbered pages 

\TitlePage 

\doublespacing

\begin{abstract} 
The single-index varying coefficient model has received much attention due to its flexibility and interpretability in recent years. This dissertation is mainly concerned with the rank-based estimation and variable selection in single-index varying coefficient models. 

In the first part of this dissertation, we consider a rank-based estimation of the index parameter and the coefficient functions for single-index varying coefficient model. The consistency and asymptotic normality of the proposed estimators are established.  An extensive Monte-Carlo simulation study demonstrates the robustness and the efficiency of the proposed estimators compared to the least squares estimators. The rank-based approach was motivated by a problem from fisheries ecology where it is used to provide accurate estimates of interspecies dependence along an environmental gradient. We use a real data example to show that the classical approach is highly affected by outliers in response space but not the rank-based method we proposed in this dissertation.

The second part of this dissertation is based on variable selection method for single-index varying coefficient model. We develop a LASSO-type rank-based variable selection procedure to select and estimate coefficient functions. A Monte-Carlo simulation study shows that the proposed method is highly robust and efficient compared to least squares type approaches. Our method can be easily applied to single-index and varying coefficient models since they are special cases of single-index varying coefficient model.
\end{abstract}

\begin{acknowledgments}
This dissertation would not have been possible without the support of many people.

First of all, I wish to take this opportunity to express my sincere gratitude to my supervisor, Dr. Asheber Abebe who was abundantly helpful and offered invaluable assistance, support and guidance. I benefit so much from his deep insight into statistics, his professionalism and his constant encouragement. I feel extremely lucky to have him as my thesis advisor.

I am also deeply grateful to my Co-advisor Dr. Nedret Billor for her invaluable advice and helpful comments. Deepest gratitude are also due to my members of the supervisory committee, Drs. Peng Zeng, Guanqun Cao and Huybrechts F.Bindele without whose knowledge and assistance this study would not have been successful.

I wish to thank all professors who taught me in the Mathematics and Statistics department for their excellent lectures, particularly, Drs. Asheber Abebe, Nedret Billor, Peng Zeng, Ming Liao and Jessica McDonald.

I appreciate the friendship with my classmates and friends, including Dr. Yi Xu, Dr. Zhefeng Jiang, Dr. Zhifeng He, Xuyu Wang, Kefan Xiao, Yu Wang, Mingjie Feng, Ninkai Tang, Dr. Hao Wu, Chi Xu, Hannah Correia, Dr. Jessica Busler, and many others. 

Finally, I would like to dedicate this thesis to my parents for their boundless love.
\end{acknowledgments}

\begin{singlespace}
	\tableofcontents
	\clearpage
	\listoffigures
	\clearpage
	\listoftables
\end{singlespace}

% \printnomenclature[0.5in] %used for the List of Abbreviations
\end{romanpages}        % All done with roman-numbered pages

\normalem       % Make italics the default for \em

\doublespacing

