---
title: ""
author: ""
output: 
  html_document:
    highlight: null
---

These notes were compiled by Ben Staton in helping to remember how to do some standard syntax for $\LaTeX$.

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = F)
```

## BibTeX containers

More details [here](http://bib-it.sourceforge.net/help/fieldsAndEntryTypes.php#inproceedings)

### Articles

```{r}
# raw
@article{,
  author = {},
  year = {},
  title = {},
  journal = {},
  volume = {},
  number = {},
  pages = {--},
  url = {},
  doi = {}
}

# example
@article{staton-etal-2017,
  author = {Benjamin A. Staton and Matthew J. Catalano and Steven J. Fleischman},
  year = {2017},
  title = {From Sequential to Integrated Bayesian Analyses: Exploring the Continuum with a Pacific Salmon Spawner-Recruit Model},
  journal = {Fisheries Research},
  volume = {186},
  number = {1},
  pages = {237--247},
  url = {https://www.sciencedirect.com/science/article/pii/S0165783616302946},
  doi = {10.1016/j.fishres.2016.09.001}
}
```

### Whole Books

```{r}
# raw
@book {,
  author = {},
  year = {},
  title = {},
  publisher = {},
  note = {}
}

# example
@book {hilborn-walters-1992,
  author = {Ray Hilborn and Carl J. Walters},
  year = {1992},
  title = {Quantitative fisheries stock assessment: choice, dynamics, and uncertainty.},
  publisher = {Springer US},
  note = {ISBN 978-1461535980}
}
```

### Conference Proceedings

```{r}
@inproceedings{,
  author = {},
  year = {},
  title = {},
  booktitle = {},
  editor = {},
  series = {},
  volume = {},
  pages = {},
  publisher = {},
  address = {}
}
```


### Technical Reports

E.g., agency reports

```{r}
# raw
@techreport{,
  author = {},
  year = {},
  title = {},
  institution = {},
  type = {},
  number = {},
  address = {},
  url = {}
}

# example
@techreport{smith-liller-2017,
  author = {Nicholas J. Smith and Zachary W. Liller},
  year = {2017},
  title = {2017 Kuskokwim River Chinook Salmon Run Reconstruction and 2018 Forecast},
  institution = {Alaska Department of Fish and Game},
  type = {Regional Information Report},
  number = {3A18-02},
  address = {Anchorage, AK},
  url = {http://www.adfg.alaska.gov/FedAidPDFs/RIR.3A.2018.02.pdf}
}
```


### 

## $\LaTeX$ containers

### Figures

To insert a figure with a caption

```{r}
\begin{figure}
  \centering
  \includegraphics{path/to/image.png}
  \caption{}
  \label{fig:}
\end{figure}
```

### Equation-Related

Insert a numbered equation. Don't forget a tag!
```{r}
\begin{equation}
  
(\#eq:)
\end{equation}
```

Insert a centered block of equations with only one equation number.

```{r}
\begin{equation}
  \begin{split}
  first $ = eq_here \\
  sec $ = eq_here
  \end{split}
(\#eq:)
\end{equation}
```
